(Deep) Induction for GADTs

### 2021-03-30 induction-paper/GADT-induction.tex
Section: General case

%Outline general heuristics for this process, generalizing process in deep induction paper.
%\begin{itemize}
%    \item Lifting for GADT (possibly a restricted class?)
%    \item Deep induction rule
%    \item Prove deep induction (define term)
%\end{itemize}

TODO:
\begin{itemize}
\item explain overline is vector
\end{itemize}

@@

Consider a generic GADT
\begin{equation}\label{eq:gadt}
\begin{array}{l}
\mathsf{data\, G\,(\ol{a : Set}) : Set\,where}\\
\mathsf{\;\;\;\;\;\;\;\;C_i\, :\, F_i\,G\,\ol{b} \to G (\ol{K_i\,\ol{b}})}
\end{array}
\end{equation}
with a finite number of constructors $\mathsf{C_i}$,
and where $\mathsf{F_i}$ has type $\mathsf{(Set^{\lvert\ol{a}\rvert} \to Set) \to Set^{\lvert\ol{b}\rvert} \to Set}$
and each $\mathsf{K_i}$ has type $\mathsf{Set^{\lvert\ol{b}\rvert} \to Set}$.
Each constructor $\mathsf{C_i}$ can be rewritten as
\[
\mathsf{C_i : \exists (\ol{b : Set}) \to \ol{Equal\,a\,(K_i\,\ol{b})} \to F_i\,G\,\ol{b} \to G\,\ol{a}}
\]
which is the form we shall use from now on.

We define the lifting of $\mathsf{G}$
\[
\mathsf{G^{\wedge} : \forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set}
\]
as
\[
\mathsf{G^{\wedge}\,\ol{a}\,\ol{Q_a}\,(C_i\,\ol{b}\,\ol{e_a}\,x)
= \exists (\ol{Q_b : b \to Set})
\to \ol{Equal^{\wedge}\,a\,(K_i\,\ol{b})\,Q_a\,(K_i^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}
\times F_i^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,{\ol{Q_b}}\,x}
\]
where $\mathsf{Q_a : a \to Set}$, $\mathsf{e_a : Equal\,a\,(K_i\,\ol{b})}$ and $\mathsf{x : F_i\,G\,\ol{b}}$,
and the liftings
\begin{multline*}
\mathsf{F_i^{\wedge} : \forall (G : Set^{\lvert\ol{a}\rvert} \to Set) (\ol{b : Set})
\to (\forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set)} \\
\mathsf{\to (\ol{b \to Set})
\to F_i\,G\,\ol{b} \to Set}
\end{multline*}
and
\[
\mathsf{K_i^{\wedge} : \forall (\ol{b : Set}) \to (\ol{b \to Set}) \to K_i\,\ol{b} \to Set}
\]
are defined by induction on the structure of the type.

Define
\[
\mathsf{PredMap : \forall (a : Set) \to (a \to Set) \to (a \to Set) \to Set}
\]
as $\mathsf{PredMap\,a\,Q_1\,Q_2 = \forall (x : a) \to Q_1\,x \to Q_2\,x}$

%  -- Lifting also acts on morphisms between predicates with the same carrier
%  {-# TERMINATING #-}
%  Bush^-map : ∀ {l} {X : Set} {Q1 Q2 : X → Set l}
%              → Pred-map {l} Q1 Q2 → Pred-map (Bush^ Q1) (Bush^ Q2)
%  Bush^-map Q12 Bnil q = bigtt
%  Bush^-map Q12 (Bcons x b) (y , bby) =
%                Q12 x y , Bush^-map (Bush^-map Q12) b bby
Define
\[
\mathsf{GLMap : \forall (a : Set) (Q_1\,Q_2 : a \to Set) \to PredMap\,a\,Q_1\,Q_2 \to PredMap\,(G\,a)\,(G^{\wedge}\,a\,Q_1)\,(G^{\wedge}\,a\,Q_2)}
\]
as
\[
\mathsf{
GLMap\,\ol{a}\,\ol{Q_1}\,\ol{Q_2}\,\ol{M}\,(C_i\,\ol{b}\,\ol{e_a}\,x)(\ol{Q_b}, \ol{e'_a}, l)
= \ol{M}
}
\]
%\begin{gather*}
%\end{gather*}

Associate the type
\begin{multline}
\mathsf{PC_i = \forall (\ol{a : Set}) (\ol{b : Set}) (\ol{Q_a : a \to Set}) (\ol{Q_b : b \to Set}) (\ol{e_a : Equal\,a\,(K_i\,\ol{b})}) (x : F_i\,G\,\ol{b})} \\
\mathsf{\to \ol{Equal^{\wedge}\,a\,(K_i\,\ol{b})\,Q_a\,(K_i^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}
	\to F_i^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x
	\to P\,\ol{a}\,\ol{Q_a}\,(C_i\,\ol{b}\,\ol{e_a}\,x)}
\end{multline}
to each constructor $\mathsf{C_i}$.

Define the induction rule for $\mathsf{G}$
\begin{multline}
\mathsf{DIG : \forall (P : \forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set)} \\
\mathsf{\to \bigtimes_i PC_i
\to \forall (\ol{a : Set})(\ol{Q_a : a \to Set})(x : G\,\ol{a}) \to G^{\wedge}\,\ol{a}\,\ol{Q_a}\,x \to P\,\ol{a}\,\ol{Q_a}\,x}
\end{multline}
as
\[
\mathsf{DIG\,P\,\ol{pc_i}\,\ol{a}\,\ol{Q_a}\,(C_i\,\ol{b}\,\ol{e_a}\,x)\,(\ol{Q_b}, \ol{e'_a}, l)
= pc_i\,\ol{a}\,\ol{b}\,\ol{Q_a}\,\ol{Q_b}\,\ol{e_a}\,x\,\ol{e'_a}\,\ol{p}}
\]
where
\[
\mathsf{p = ??? : F_i^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}
\]

\begin{lemma}
For every $\mathsf{F, G, \ol{b}, P, \ol{Q_b}, x}$, there exists a function
\[
\mathsf{p : F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x \to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}
\]
\end{lemma}
\begin{proof}
By structural induction on $\mathsf{F}$:
\begin{itemize}
\item Case $\mathsf{F = F_1 \times F_2}$.
Thus, $\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x = F_1^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x \times F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}$.
By inductive hypothesis, there exist functions
$\mathsf{p_1 : F_1^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x_1 \to ??? \to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x_1}$ and
$\mathsf{p_2 : F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x_2 \to ??? \to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x_2}$.
%%%
\item Case $\mathsf{F\,G\,\ol{b} = G(\ol{F_a\,\ol{b}})}$.
Thus, $\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b} = P (\ol{F_a\,\ol{b}}) (\ol{F_a^{\wedge}\,\ol{b}\,\ol{Q_b}})}$.
So, $\mathsf{p : P (\ol{F_a\,\ol{b}}) (\ol{F_a^{\wedge}\,\ol{b}\,\ol{Q_b}}) x}$ is
\[
\mathsf{p = DIG\,P\,\ol{pc_i}\,(\ol{F_a\,\ol{b}})\,(\ol{F_a^{\wedge}\,\ol{b}\,\ol{Q_b}})\,x\,l}
\]
Note: No nesting!
\end{itemize}
\end{proof}

@@

\begin{itemize}
\item reference
\end{itemize}

>>>>

@@

Consider a generic GADT
\begin{equation}\label{eq:gadt}
\begin{array}{l}
\mathsf{data\, G\,(\ol{a : Set}) : Set\,where}\\
\mathsf{\;\;\;\;\;\;\;\;C\, :\, F\,G\,\ol{b} \to G (\ol{K\,\ol{b}})}
\end{array}
\end{equation}
which, for ease of notation,
we assume to have only one constructor $\mathsf{C}$.
Extending the argument to the case of a GADT with multiple constructors presents no difficulty.
In the definition above, $\mathsf{F}$ has type $\mathsf{(Set^{\alpha} \to Set) \to Set^{\beta} \to Set}$
and each $\mathsf{K}$ has type $\mathsf{Set^{\beta} \to Set}$.
The overline notation denotes a finite list:
%and the vertical bars notation denotes the length of a list:
$\mathsf{G}$ depends on a finite list of types $\mathsf{\ol{a}}$ of length $\mathsf{\alpha}$,
each $\mathsf{K}$ depends on a finite list of types $\mathsf{\ol{b}}$ of length $\mathsf{\beta}$,
and notice that the length of the list $\mathsf{\ol{a}}$ matches that of the list $\mathsf{\ol{K}}$.
With that in mind, the constructor $\mathsf{C}$ can be rewritten as
\[
\mathsf{C : \exists (\ol{b : Set}) \to \ol{Equal\,a\,(K\,\ol{b})} \to F\,G\,\ol{b} \to G\,\ol{a}}
\]
which is the form we shall use from now on.

In order to state the induction rule for $\mathsf{G}$,
we first need to define its associated predicate-lifting function 
\[
\mathsf{G^{\wedge} : \forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set}
\]
as
\[
\mathsf{G^{\wedge}\,\ol{a}\,\ol{Q_a}\,(C\,\ol{b}\,\ol{e_a}\,x)
= \exists (\ol{Q_b : b \to Set})
\to \ol{Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}
\times F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,{\ol{Q_b}}\,x}
\]
where $\mathsf{Q_a : a \to Set}$, $\mathsf{e_a : Equal\,a\,(K\,\ol{b})}$ and $\mathsf{x : F\,G\,\ol{b}}$.
Notice that we need a type $\mathsf{Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}$
for each type $\mathsf{a}$.
We also assume to have liftings for $\mathsf{F}$,
\begin{multline*}
\mathsf{F^{\wedge} : \forall (G : Set^{\alpha} \to Set) (\ol{b : Set})
\to (\forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set)} \\
\mathsf{\to (\ol{b \to Set})
\to F\,G\,\ol{b} \to Set}
\end{multline*}
and for each  $\mathsf{K}$,
\[
\mathsf{K^{\wedge} : \forall (\ol{b : Set}) \to (\ol{b \to Set}) \to K\,\ol{b} \to Set}
\]
as we could not possibly define an induction rule otherwise.
A way to concretely define the predicate-lifting function for a type is to proceed by induction on the structure of the type.
% on the structure of $\mathsf{F}$ and $\mathsf{K}$ respectively.

%Associate the type
%\begin{multline}
%\mathsf{PC = \forall (\ol{a : Set}) (\ol{b : Set}) (\ol{Q_a : a \to Set}) (\ol{Q_b : b \to Set}) (\ol{e_a : Equal\,a\,(K\,\ol{b})}) (x : F\,G\,\ol{b})} \\
%\mathsf{\to \ol{Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}
%	\to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x
%	\to P\,\ol{a}\,\ol{Q_a}\,(C\,\ol{b}\,\ol{e_a}\,x)}
%\end{multline}
%to the constructor $\mathsf{C}$.

Define the induction rule for $\mathsf{G}$
\begin{multline}
\mathsf{DIG : \forall (P : \forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set)} \\
\mathsf{\to \Big(
\forall (\ol{a : Set}) (\ol{b : Set}) (\ol{Q_a : a \to Set}) (\ol{Q_b : b \to Set}) (\ol{e_a : Equal\,a\,(K\,\ol{b})}) (x : F\,G\,\ol{b})} \\
\mathsf{\to \ol{Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}
	\to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x
	\to P\,\ol{a}\,\ol{Q_a}\,(C\,\ol{b}\,\ol{e_a}\,x)
\Big)} \\
\mathsf{\to \forall (\ol{a : Set})(\ol{Q_a : a \to Set})(x : G\,\ol{a}) \to G^{\wedge}\,\ol{a}\,\ol{Q_a}\,x \to P\,\ol{a}\,\ol{Q_a}\,x}
\end{multline}
as
\[
\mathsf{DIG\,P\,PC\,\ol{a}\,\ol{Q_a}\,(C\,\ol{b}\,\ol{e_a}\,x)\,(\ol{Q_b}, \ol{E_a}, L_F)
= PC\,\ol{a}\,\ol{b}\,\ol{Q_a}\,\ol{Q_b}\,\ol{e_a}\,x\,\ol{E_a}\,(p\,x\,L_F)}
\]
where
%$\mathsf{\ol{pc_i : PC_i}}$,
%$\mathsf{\ol{Q_a : a \to Set}}$,
we have $\mathsf{e_a : Equal\,a\,(K\,\ol{b})}$ for each $\mathsf{a}$,
$\mathsf{x : F\,G\,\ol{b}}$,
$\mathsf{Q_b : b \to Set}$ for each $\mathsf{b}$ ,
$\mathsf{E_a : Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}$  for each $\mathsf{a}$,
$\mathsf{L_F : F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,{\ol{Q_b}}\,x}$,
and
\[
\mathsf{p : PredMap\,(F\,G\,\ol{b})\,(F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}) (F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b})}
\]
is defined by structural induction on $\mathsf{F}$ as follows:
\begin{itemize}
\item Case $\mathsf{F = F_1 \times F_2}$ where $\mathsf{F_1}$ and $\mathsf{F_2}$ have the same type as $\mathsf{F}$.
We have that
\[
\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}
= Pair^{\wedge}(F_1\,G\,\ol{b})(F_2\,G\,\ol{b})(F_1^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}) (F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b})}
\]
By inductive hypothesis, there exist morphisms of predicates
\begin{gather*}
\mathsf{p_1 : PredMap\,(F_1\,G\,\ol{b})\,((F_1)^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}) ((F_1)^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b})} \\
\mathsf{p_2 : PredMap\,(F_2\,G\,\ol{b})\,((F_2)^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}) ((F_2)^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b})}
\end{gather*}
Thus, we define $\mathsf{p (x_1, x_2) (L_1, L_2) = (p_1\,x_1\,L_1, p_2\,x_2\,L_2)}$ 
for 
$\mathsf{x_1 : F_1\,G\,\ol{b}}$,
$\mathsf{x_2 : F_2\,G\,\ol{b}}$,
$\mathsf{L_1 : F_1^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x_1}$
and $\mathsf{L_2 : F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x_2}$.
%%%
\item Case $\mathsf{F = F_1 + F_2}$ where $\mathsf{F_1}$ and $\mathsf{F_2}$ have the same type as $\mathsf{F}$.
Analogous to case $\mathsf{F = F_1 \times F_2}$.
%%%
\item Case $\mathsf{F = F_1 \to F_2}$
where $\mathsf{F_1}$ and $\mathsf{F_2}$ have the same type as $\mathsf{F}$.
%where $\mathsf{F_1}$ does not contain the recursive variable,
%i.e., $\mathsf{F_a : Set^\beta \to Set}$,
%and $\mathsf{F_2}$ has the same type as $\mathsf{F}$.
We have that
\begin{gather*}
\mathsf{F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x = \forall (y : F_1\,G\,\ol{b}) \to F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,(x\,y)} \\
%F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,x \to F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x} \\
\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x = \forall (y : F_1\,G\,\ol{b}) \to F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,(x\,y)}
%\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x = F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,x \to F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}
\end{gather*}
By inductive hypothesis, there exist a morphism of predicates
\[
\mathsf{p_2 : PredMap\,(F_2\,G\,\ol{b}) (F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}) (F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b})}
\]
Thus, we define $\mathsf{p\,x\,L_F\,y = p_2 (x\,y) (L_F\,y)}$.
{\color{blue} Revise this case!}
%Thus, we define $\mathsf{p\,x\,L_F\,L_1 = p_2\,(L_F\,L_1)}$
%for $\mathsf{L_1 : F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,x}$.
%Notice that $\mathsf{F_1}$ does not contain the recursive variable.
%That is a necessary restriction, as the proof relies on
%$\mathsf{F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x}$
%and $\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}$
%having the same domain.
%%%
\item Case $\mathsf{F\,G\,\ol{b} = G(\ol{F_a\,\ol{b}})}$
where none of the $\mathsf{F_a}$ contains the recursive variable,
i.e., $\mathsf{F_a : Set^\beta \to Set}$
for each $\mathsf{a}$.
Thus, $\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b} = P (\ol{F_a\,\ol{b}}) (\ol{F_a^{\wedge}\,\ol{b}\,\ol{Q_b}})}$.
So, $\mathsf{p}$ is defined as
\[
\mathsf{p = DIG\,P\,PC\,(\ol{F_a\,\ol{b}})\,(\ol{F_a^{\wedge}\,\ol{b}\,\ol{Q_b}})}
\]
%for $\mathsf{L_F : F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x}$.
Notice that the fact that the types $\mathsf{F_a}$ do not contain the recursive variable
means that $\mathsf{G}$ does not feature nesting.
As mentioned before, that is a fundamental assumption to make the argument work.
Why the argument fails in presence of truly nested GADTs will be shown in the next section.
%%%
\item Case $\mathsf{F\,G\,\ol{b} = H\,\ol{b}}$
where $\mathsf{H}$ is a type constructor not containing the recursive variable,
i.e., $\mathsf{H : Set^\beta \to Set}$.
In such case, $\mathsf{p : PredMap (H\,\ol{b}) (H^{\wedge}\,\ol{b}\,\ol{Q_b}) (H^{\wedge}\,\ol{b}\,\ol{Q_b})}$
is just the identity morphism of predicates.
%%%
\item Case $\mathsf{F\,G\,\ol{b} = H\,(\ol{F_c\,G\,\ol{b}})}$
where $\mathsf{H}$ is a type constructor not containing the recursive variable,
i.e., $\mathsf{H : Set^\gamma \to Set}$,
and $\mathsf{F_c}$ has the same type as $\mathsf{F}$, for every $\mathsf{c = 1 \dots \gamma}$.
Moreover, we assume that $\mathsf{H}$ has an associated predicate-lifting function,
\[
\mathsf{H^{\wedge} : \forall (\ol{c : Set}) \to (\ol{c \to Set}) \to H\,\ol{c} \to Set}
\]
and that this predicate-lifting function has a map function
\[
\mathsf{HLMap : \forall (\ol{c : Set}) (\ol{Q_c\;Q_c' : c \to Set}) \to PredMap\,\ol{c}\,\ol{Q_c}\,\ol{Q_c'}
\to PredMap\,(H\,\ol{c})\,(H^{\wedge}\,\ol{c}\,\ol{Q_c})\,(H^{\wedge}\,\ol{c}\,\ol{Q_c'})}
\]
That means that $\mathsf{H}$ cannot be a GADT, as those do not have functorial semantics~\cite{HaskellPaper},
but it can be an ADT or even a nested type~\cite{jp19,jgj21}.
Thus,
\[
\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b} = H^{\wedge} (\ol{F_c\,G\,\ol{b}}) (\ol{F_c^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}})}
\]
By induction hypothesis, there is a morphism of predicates
\[
\mathsf{p_c : PredMap\,(F_c\,G\,\ol{b}) (F_c^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}) (F_c^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b})}
\]
for every $\mathsf{c = 1 \dots \gamma}$.
So, $\mathsf{p}$ is defined as
\[
\mathsf{p
= HLMap\,(\ol{F_c\,G\,\ol{b}}) (\ol{F_c^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}}) (\ol{F_c^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}})\,\ol{p_c}}
\]

\end{itemize}
Notice that, by proceeding by structural induction as above,
we have implicitly defined a grammar for GADTs.
In particular, we allow $\mathsf{F}$ to feature sums, products, arrow types and (limited) type application.

%\subsection{General case}
%%Outline general heuristics for this process, generalizing process in deep induction paper.
%%\begin{itemize}
%%    \item Lifting for GADT (possibly a restricted class?)
%%    \item Deep induction rule
%%    \item Prove deep induction (define term)
%%\end{itemize}
%TODO:
%\begin{itemize}
%\item explain overline is vector
%\end{itemize}
%In this section we will generalize the approach taken in the previous examples
%and provide a general framework to derive induction rules for GADTs.
%We restrict our scope to GADTs not featuring nesting,
%in the sense that their constructors contain no nested calls of the recursive variable,
%as truly nested types do.
%The reason for this restriction will be discussed in Section~\ref{sec:GADT-nested}.
%Consider a generic GADT
%\begin{equation}\label{eq:gadt}
%\begin{array}{l}
%\mathsf{data\, G\,(\ol{a : Set}) : Set\,where}\\
%\mathsf{\;\;\;\;\;\;\;\;C_i\, :\, F_i\,G\,\ol{b} \to G (\ol{K_i\,\ol{b}})}
%\end{array}
%\end{equation}
%with a finite number of constructors $\mathsf{C_i}$,
%and where $\mathsf{F_i}$ has type $\mathsf{(Set^{\lvert\ol{a}\rvert} \to Set) \to Set^{\lvert\ol{b}\rvert} \to Set}$
%and each $\mathsf{K_i}$ has type $\mathsf{Set^{\lvert\ol{b}\rvert} \to Set}$.
%The overline notation denotes a finite list:
%$\mathsf{G}$ depends on a finite list of types $\mathsf{a}$,
%each $\mathsf{K_i}$ depends on a finite list of types $\mathsf{b}$,
%and notice that the length of the list of $\mathsf{a}$ matches that of the list of $\mathsf{K_i}$.
%With that in mind, each constructor $\mathsf{C_i}$ can be rewritten as
%\mathsf{C_i : \exists (\ol{b : Set}) \to \ol{Equal\,a\,(K_i\,\ol{b})} \to F_i\,G\,\ol{b} \to G\,\ol{a}}
%which is the form we shall use from now on.
%We define the lifting of $\mathsf{G}$
%\mathsf{G^{\wedge} : \forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set}
%\mathsf{G^{\wedge}\,\ol{a}\,\ol{Q_a}\,(C_i\,\ol{b}\,\ol{e_a}\,x)
%= \exists (\ol{Q_b : b \to Set})
%\to \ol{Equal^{\wedge}\,a\,(K_i\,\ol{b})\,Q_a\,(K_i^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}
%\times F_i^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,{\ol{Q_b}}\,x}
%where $\mathsf{Q_a : a \to Set}$, $\mathsf{e_a : Equal\,a\,(K_i\,\ol{b})}$ and $\mathsf{x : F_i\,G\,\ol{b}}$,
%and the liftings
%\mathsf{F_i^{\wedge} : \forall (G : Set^{\lvert\ol{a}\rvert} \to Set) (\ol{b : Set})
%\to (\forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set)} \\
%\mathsf{\to (\ol{b \to Set})
%\to F_i\,G\,\ol{b} \to Set}
%\mathsf{K_i^{\wedge} : \forall (\ol{b : Set}) \to (\ol{b \to Set}) \to K_i\,\ol{b} \to Set}
%are defined by induction on the structure of the type.
%Associate the type
%\mathsf{PC_i = \forall (\ol{a : Set}) (\ol{b : Set}) (\ol{Q_a : a \to Set}) (\ol{Q_b : b \to Set}) (\ol{e_a : Equal\,a\,(K_i\,\ol{b})}) (x : F_i\,G\,\ol{b})} \\
%\mathsf{\to \ol{Equal^{\wedge}\,a\,(K_i\,\ol{b})\,Q_a\,(K_i^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}
%	\to F_i^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x
%	\to P\,\ol{a}\,\ol{Q_a}\,(C_i\,\ol{b}\,\ol{e_a}\,x)}
%to each constructor $\mathsf{C_i}$.
%Define the induction rule for $\mathsf{G}$
%\mathsf{DIG : \forall (P : \forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set)} \\
%\mathsf{\to \bigtimes_i PC_i
%\to \forall (\ol{a : Set})(\ol{Q_a : a \to Set})(x : G\,\ol{a}) \to G^{\wedge}\,\ol{a}\,\ol{Q_a}\,x \to P\,\ol{a}\,\ol{Q_a}\,x}
%\mathsf{DIG\,P\,\ol{pc_i}\,\ol{a}\,\ol{Q_a}\,(C_i\,\ol{b}\,\ol{e_a}\,x)\,(\ol{Q_b}, \ol{E_a}, L_F)
%= pc_i\,\ol{a}\,\ol{b}\,\ol{Q_a}\,\ol{Q_b}\,\ol{e_a}\,x\,\ol{E_a}\,(p\,L_F)}
%where $\mathsf{\ol{pc_i : PC_i}}$,
%$\mathsf{\ol{Q_a : a \to Set}}$,
%$\mathsf{\ol{e_a : Equal\,a\,(K_i\,\ol{b})}}$,
%$\mathsf{x : F_i\,G\,\ol{b}}$,
%$\mathsf{\ol{Q_b : b \to Set}}$,
%$\mathsf{\ol{E_a : Equal^{\wedge}\,a\,(K_i\,\ol{b})\,Q_a\,(K_i^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}}$,
%$\mathsf{L_F : F_i^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,{\ol{Q_b}}\,x}$,
%and $\mathsf{p}$ is defined in the following lemma.
%\begin{lemma}
%For every $\mathsf{F, G, \ol{b}, P, \ol{Q_b}, x}$, there exists a function
%\mathsf{p : F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x \to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}
%\end{lemma}
%\begin{proof}
%By structural induction on $\mathsf{F}$:
%\begin{itemize}
%\item Case $\mathsf{F = F_1 \times F_2}$.
%We have that $\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x = F_1^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x \times F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}$.
%By inductive hypothesis, there exist functions
%$\mathsf{p_1 : F_1^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x_1 \to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x_1}$ and
%$\mathsf{p_2 : F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x_2 \to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x_2}$.
%Thus, we define $\mathsf{p (L_1, L_2) = (p_1\,L_1) \times (p_2\,L_2)}$
%%%%
%\item Case $\mathsf{F = F_1 + F_2}$. Analogous to case $\mathsf{F = F_1 \times F_2}$.
%%%%
%\item Case $\mathsf{F = F_1 \to F_2}$ where $\mathsf{F_1}$ does not contain the recursive variable.
%We have that
%\mathsf{F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x = F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,x \to F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x} \\
%\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x = F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,x \to F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}
%By inductive hypothesis, there exist a function
%$\mathsf{p_2 : F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x_2 \to F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x_2}$.
%Thus, we define $\mathsf{p\,L\,l = p_2\,(L\,l)}$ for
%$\mathsf{L : F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}$
%and $\mathsf{l : F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,x}$.
%{\color{blue} TODO: explain why $\mathsf{F_1}$ does not contain the recursive variable.}
%%%%
%\item Case $\mathsf{F\,G\,\ol{b} = G(\ol{F_a\,\ol{b}})}$.
%Thus, $\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b} = P (\ol{F_a\,\ol{b}}) (\ol{F_a^{\wedge}\,\ol{b}\,\ol{Q_b}})}$.
%So, $\mathsf{p}$ is defined as
%\mathsf{p\,L_F = DIG\,P\,\ol{pc_i}\,(\ol{F_a\,\ol{b}})\,(\ol{F_a^{\wedge}\,\ol{b}\,\ol{Q_b}})\,x\,L_F}
%Note: No nesting!
%\end{itemize}
%\end{proof}

@@

\begin{itemize}
\item reference (correctly) Haskell Symposium paper
\end{itemize}
